\section*{Foreword}
Originally inspired by the Nordic Collegiate Programming Contest (NCPC),
it has been held at NTNU every spring since 2007. The format is a
five-hour contest with competing teams consisting of one, two or three
contestants. A team of volunteer judges write the problems and answer
clarification requests during the contest, while another team hands out
balloons for each solved problem. Usually a rather hectic affair, it is
extremely important that everything is well prepared. The number of
teams is often more than 100, with the record being 162 teams in 2011

The contest system that verifies solutions is at the heart of the
contest when it is in progress, and needs to be working perfectly at
all times. The system must handle several submissions per second, while
verifying that each one is correct and runs within the set resource
limits. Submissions must show up on the high score list, and when
problems are solved the team handing out balloons must be notified. In
addition to this there were a lot of other functional requirements
having to do with the bureaucracy of organizing the contest

A requirement was that new features could be easily added in the future,
and the code was written with this in mind. The project will now become
open source, and all programming contest enthusiasts will soon be able
to request and implement their desired features

All aspects of this project have been pleasing and delightful for us.
The team has exceeded all our expectations and their system will be
used for years to come.

\hfill -- \textit{Christian Chavez, IDI Open Manager}

\pagebreak

\section*{Preface}

IT2901 is the most prestigious course for bachelor students studying
informatics. This motivated us to give the best of our self. This report is
written spring 2014 by six students at NTNU, Norwegian University of Science
and Technology, Department of Computer and Information Science, IDI. It has
been a great and educational journey. 

For the past few months, we have taught ourselves a new programming language
and framework and used advanced development tools - while tackling many social
and technical conflicts.  We’ve proven how ``Ambition is a dream with a V8
engine'', as Elvis Presley once said. 

The group would like to thank our eager customers, Christian Chavez, Finn
Inderhaug Holme and Christian Neverdal Jonassen for their time to meet us and
provide constructive feedback. We also owe a big  thanks to our friends and
family, for understanding our choice to work during easter. Big thank you to
our supervisor, Hong Guo, for constructive criticism and reflections; without
which, we would not ascertain the peak of our own potential.
