\chapter{Project Management}
This chapter will go through the different project roles we deemed
important. We will explain our development method, which tools we use
and give an overview of how we planned the project. Furthermore, in
section~\ref{section:WBS}. We also provide a structured overview of how we organized
our time. 

\section{Project Roles}
We wanted to ensure that all developers had an even workload and
experience in all components of our project. We
maintained a flat organizational structure where all decisions were
made in groups. No member would work alone on a task for a longer
period of time. Some tasks and delegations were easier to assign only once.

The most central role is that of the scrum master. The role mainly
consists of setting up meeting agendas and keeping control of what
team members are working on. In addition, the scrum master should act
as a buffer between the team and other distractions. The scrum master had a
casting vote whenever there was a disagreement. The group elected Haakon to be
scrum master because of his well established authority and organization.

We also assigned the role of a transcriptionist. His job consists of
writing a short summary of every meeting, and making this available to
the rest of the group. This includes meetings with the customer and
supervisor. This job was performed by Anders, who volunteered for the
position. We assigned Håkon to be customer contact, and
Tino as responsible for room reservations.

\section{Development Method (Scrum)}
Scrum focuses on having daily meetings, and constantly adjusting to
changes by iterative development. This makes it easier to predict and
to adjust for problems that may occur. It was hard to predict what
would happen in our project, therefore our sprints were short, lasting
at most two weeks. The transition between two sprints was done during a
prolonged meeting on Wednesdays. During this meeting we evaluated the
latest sprint and planned the upcoming one. Every team member were
requested to say three good things and three bad things regarding the
last sprint. This was followed by a discussion of how to plan the next
sprint better. Lastly we showed what had been completed, to the other
members of the group, before setting up the next sprint. Scrum also
focuses on having finished versions of the systems on each iteration,
and to finish all packages in the given iteration.
In order to take advantage of the best in everyone's
abilities we worked in pairs where this was efficient. Working in pairs
is common in agile development. This was to improve code quality and
reduce errors.
%\footnote{[Page 2}
%href{http://www.cs.pomona.edu/classes/cs121/supp/williams\_prpgm.pdf}{http}

\section{Tools/Framework}
The customer wanted our end product to be easy to maintain for future
developers. Therefore we have chosen tools that are well known and easy
to learn. Some of the most important are:

\begin{itemize}
    \item Django, a framework written in Python.
    \item VIM and Eclipse for editing
    \item Google Drive and latex for documentation
    \item Git as version control, with github as hosting service
    \item Email lists, IRC and Facebook for communication
    \item Bootstrap and Grappelli for user interface design
\end{itemize}
A lot of different tools were considered for this system. A full list of all
tools and frameworks used and considered can be viewed in appendix
\textit{Tools and Frameworks.}

\section{Project-Level Planning}
After our initial requirements elicitation we began to plan our
development process. The purpose of the plan was to verify that we had
enough time to complete the requirements, and to avoid unforeseen
risks. This section will present the various components we introduced
to structurize the project.

\subsection{Work Breakdown Structure}\label{section:WBS}
WBS is a decomposition of the project into phases, deliverables and work
packages. Each package was further broken down into different tasks.
The benefits from the WBS are as follows:
\begin{itemize}
    \item Planning out the entire process prevents bottlenecks
    \item Clearly defining the scope of a package prevents excess or
        insufficient time usage
    \item It is easy for supervisors and other parties to evaluate and
        understand our process
\end{itemize}

Table~\ref{table:WBS} shows the work breakdown structure created. These high-level
packages were later broken down into activities, which are in the
product backlog, see appendix~\ref{appendix:productBacklog}

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
\scriptsize
\begin{longtable}{|p{.8\textwidth}|}
    \caption[test]{Work breakdown structure} \label{table:WBS} \\
\hline
\begin{enumerate}[nosep]
    \itemsep0em 
        \item Project management
        \begin{enumerate}[nosep]
    \itemsep0em 
            \item Write timesheet template
            \item Look at the reflection notes
            \item Meetings
            \begin{enumerate}[label*=\arabic*.]
    \itemsep0em 
                \item Internal
                \item Customer
                \item supervisor
            \end{enumerate}
            \item  Report
            \begin{enumerate}[label*=\arabic*.]
    \itemsep0em 
                \item Preliminary version
                \item Mid-semester version
                \item Final version
            \end{enumerate}
            \item Risk assessment
            \item WBS
            \item Status report
            \item Activity plans
        \end{enumerate}

        \item Pre-study        
        \begin{enumerate}[nosep]
    \itemsep0em 
            \item Install and learn tools
            \item Learn language/framework
            \item Course
        \end{enumerate}
        \item Design
            \begin{enumerate}[nosep]
    \itemsep0em 
            \item Requirement Specification
            \begin{enumerate}[label*=\arabic*.]
    \itemsep0em 
                \item Functional
                \item Non-functional
            \end{enumerate}
            \item  System architecture
            \item Database modeling
            \item User Interface
                \begin{enumerate}[label*=\arabic*.]
    \itemsep0em 
                \item Prototyping
                \item Usability Testing
            \end{enumerate}
            \item  Admin interface
        \end{enumerate}

        \item Development
        \begin{enumerate}[nosep]
    \itemsep0em 
            \item  Backend
            \begin{enumerate}[label*=\arabic*.]
                \item Execution-node(s)
                    \begin{enumerate}[label*=\arabic*.]
                    \itemsep0em 
                    \item Web-page
                \end{enumerate}
            \end{enumerate}
            \item User management
            \begin{enumerate}[label*=\arabic*.]
                \itemsep0em 
                \item User
                \item Usergroups
                \item Team management
            \end{enumerate}
            \item  Statistics
            \item Contest management 
            \item Clarification system
            \item Balloons system
            \item Unit testing
        \end{enumerate}

        \item Testing
        \begin{enumerate}[nosep]
    \itemsep0em 
            \item  User-test
            \item System-test  
            \item Final test
        \end{enumerate}

        \item Implementation
        \begin{enumerate}[nosep]
    \itemsep0em 
            \item  Deploy to production
            \item Installation
            \item Turn in to stakeholder
        \end{enumerate}

        \item Implementation
        \begin{enumerate}[nosep]
    \itemsep0em 
            \item Verify
            \item Document
        \end{enumerate}
    \end{enumerate} \\
\hline
\end{longtable}
\normalsize

We also created a gantt chart. Here, each package was assigned an
estimated time period, over how long time we expected to use. For ease
of comprehension, not every package was included from the WBS.\ The
gantt chart is shown in figure~\ref{gantt}
\newline
\begin{longtable}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
\caption{Gantt chart} \label{gantt} \\
\hline
\textbf{WP Name} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
\hline
Project management & \cellcolor{ForestGreen} &\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen}&\cellcolor{ForestGreen} \\
\hline
WBS &\cellcolor{YellowGreen}&\cellcolor{YellowGreen}&&&&&&&&&&&&& \\
\hline
Pre-study
&\cellcolor{yellow}&\cellcolor{yellow}&\cellcolor{yellow}&\cellcolor{yellow}&&&&&&&&&&&
\\ \hline
Install and learn tools&\cellcolor{Dandelion}&\cellcolor{Dandelion}&\cellcolor{Dandelion}&\cellcolor{Dandelion}&&&&&&&&&&& \\
\hline
Learn language/framework&&\cellcolor{Dandelion}&\cellcolor{Dandelion}&\cellcolor{Dandelion}&&&&&&&&&&&  \\
\hline
Course &&\cellcolor{Dandelion}&\cellcolor{Dandelion}&&&&&&&&&&&&  \\
\hline
\textbf{Design} & &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} &\cellcolor{MidnightBlue} & & & & & \\ \hline
Requirement specification&&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&&&&&  \\
\hline
System architecture&&&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&&&&&&&&  \\
\hline
Database modelling&&&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&&&&&&&&&  \\
\hline
Tests &&&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&\cellcolor{RoyalBlue}&&&&&&&&&  \\
\hline
User-interface&&&&&\cellcolor{MidnightBlue}&\cellcolor{MidnightBlue}&\cellcolor{MidnightBlue}&\cellcolor{MidnightBlue}&\cellcolor{MidnightBlue}&&&&&&  \\
\hline
\textbf{Development}&&&&\cellcolor{Purple}&\cellcolor{Purple}&\cellcolor{Purple}&\cellcolor{Purple}&\cellcolor{Purple}&\cellcolor{Purple}&\cellcolor{Purple}&\cellcolor{Purple}&&&&  \\
\hline
Execution node&&&&&&&&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&&&&  \\
\hline
Implement single node&&&&&&&&\cellcolor{Thistle}&\cellcolor{Thistle}&\cellcolor{Thistle}&&&&&  \\
\hline
Implement several nodes&&&&&&&&&\cellcolor{Thistle}&\cellcolor{Thistle}&&&&&  \\
\hline
Content Management System&&&&&&&&&&&\cellcolor{Orchid}&&&&  \\
\hline
Front end&&&&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&&&&  \\
\hline
\textbf{Testing}&&&&&&&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}&\cellcolor{Red}  \\
\hline
Unit testing &&&&&&&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}&\cellcolor{Orchid}  \\
\hline
Integration testing&&&&&&&&&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}  \\
\hline
System test&&&&&&&&&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}&\cellcolor{Melon}  \\
\hline
\textbf{Production}&&&&&&&&&&&&&&\cellcolor{MidnightBlue}&\cellcolor{MidnightBlue}  \\
\hline
Post-implementation&&&&&&&&&&&&&&&\cellcolor{Plum} \\
\hline
\end{longtable}

The gantt chart was revised several times during the first four sprints,
mainly due to new deadlines set by the customer. 

\pagebreak
\subsection{Milestones}
\label{sec:milestones}
Throughout the project, the supervisor, customer, and the project group set
deadlines. Some of the milestones marks the completion of work packages. We
have four of these milestones, M-03, M-05, M-06 and M-07. The other milestones
represents events with deadlines that were given by the course stakeholders.
These are M-01, M-02, M-04, M-08. The group used the milestones in order
to determine if the project is on schedule and to monitor the progress.
The reader can view what requirements that were met for each milestone
in~\ref{section:functionalreq}.

\begin{longtable}{|l|p{1.3cm}|p{1.3cm}|p{1.3cm}|p{1.3cm}|p{1.3cm}|p{1.3cm}|p{1.3cm}|p{1.3cm}|}
    \caption{Milestone name and deadlines} \label{table:milestone} \\
\hline

Date & 09.02. & 09.03. & 19.03. & 19.03. & 11.04. & 26.04. & 03.05. & 30.05.\\
\hline

Week & 3 & 6 & 8 & 8 & 11 & 14 & 15 & 18\\
\hline

Name & Prelimin- ary report & Mid-semester report & First release &
Presen- tation & Beta release & IDI Open test event & IDI \ Open & Final Report\\
\hline

ID & M-01 & M-02 & M-03 & M-04 & M-05 & M-06 & M-07 & M-08\\
\hline
\end{longtable}



\begin{description}
    \item[Preliminary report M-01]
    Preliminary report is the delivery of the first version of the report.
    This was to help us get started with important aspects of the project work.

    \item[Mid-semester report M-02]
    This version of the report should present all of the analysis and most of
    the design of our system. The delivery date for the mid-semester report
    is 16th of March. We wanted to complete this one week earlier, 9th of March,
    focus on M-03.

    \item[First release M-03]
    This milestone marks the groups first delivery to the customer. In summary
    this release should make it possible for contestants to sign up for a
    competition. Three days prior to the release the group will meet up with
    the customer and overlook that all the requirements are met. This meeting
    will also act as an introduction on how to manage the system.

    \item[Presentation M-04]
    The main purpose of the presentation is for the class to share their
    experiences and learn from other groups. 

    \item[Beta-release M-05]
    The beta release should contain most of the essential features. This
    version of the program should only be a release to a selected group of
    people. 

    \item[IDI Open test event M-06]
    On April 26th there will be a test event where everybody could test the
    system. This means that leading up to this event the system should be a
    release candidate. 

    \item[IDI Open M-07]
    This is the day of the competition and the system should be a release
    version. 

    \item[Final report M-08]
    This milestone marks the final date for delivering the report and the end
    of our project. Based on feedback received from the competition the group
    might choose to implement some changes to the system. 
\end{description}

\subsection{Meetings}
Our meetings can be categorized in three categories: internal, supervisor and
customer meeting. We established some meetings rules:
\begin{itemize}
    \item All meetings follow ``the academic quarter'', meaning that the time
        of start was XX:15
    \item Members that were late had to bring a cake to the next meeting
    \item All members may at any time propose a coffee break, a proposal that
        has to be followed. 
    \item No laptop should be open during the meetings
\end{itemize}

\subsubsection{Internal Meetings}
We had three internal meetings each week. Two of which were daily scrum
meetings. These were primarily set to be on Mondays and Thursdays.
During these meetings each group member would answer three questions: 
\begin{itemize}
    \item What have you done since the last meeting?
    \item What are you planning to do until the next meeting?
    \item Do you have any problems regarding the completion of your task? 
\end{itemize}
The group would usually continue to work together after these meetings. 

On Wednesday we had longer meetings marking the end of one sprint and
the beginning of the next. This meeting would consist of a sprint
review meeting and a sprint retrospective, where we discussed
\begin{itemize}
    \item What was good/bad with the last sprint?
    \item What should we try to improve during the next sprint?
\end{itemize}
After that we held a sprint planning meeting and created a new sprint
backlog. Our official meeting structure for this meeting can be viewed in
Appendix~\ref{appendix:endOfSprint}

\subsubsection{Supervisor Meeting}
Meetings with the supervisor was generally held at a biweekly basis.
During these meetings we talked about what we had done, what we were
going to do and received feedback on what we had done. Our template
for meeting transcriptions is given in Appendix~\ref{appendix:meetingagenda}.
Before each
meeting we had to deliver status reports and activity charts. These
activity diagrams were early on replaced by sprint backlog and burndown
charts to facilitate the development process. 

\subsubsection{Customer Meeting}
Customer meetings were held whenever we felt that a certain part of the
requirements specification was unclear to us, and when we wanted approval of a
newly completed feature. Throughout the semester there were a lot of meetings.
As we never decided upon a fixed interval between customer meetings, the
frequency varied a lot. The couple of days leading up to a release date often
contained customer meetings in order to get everything right before starting on
the next release. During our periods of focusing on writing this report, the
frequency of these meetings naturally went down as the product did not
progress, and as a consequence we had little to discuss with the customer. 

\subsection{Resources}
This section contains the available resources for the project. We
intended to use a minimum of 20/25 hours per person each week, but
prepared for more work as we approached the deadline. This estimate was
later scaled up to a minimum of 25/30 two weeks before easter. During
easter, the amount of hours per week scaled up higher. 

\subsubsection{Planned Work}
Table~\ref{table:sprintoverview} shows our first initial draft of sprints. 

\begin{longtable}{|l|l|l|l|}
\caption{Initial sprint overview} \label{table:sprintoverview} \\
\hline
 Sprint & Range (week) & Days &
 Hours\\\hline
 1 & 3 - 4 & 7 & 15\\\hline
 2 & 4 - 5 & 7 & 20\\\hline
 3 & 5 - 6 & 7 & 20 \\\hline
 4 & 6 - 7 & 7 & 20 \\\hline
 5 & 7 - 8 & 7 & 20\\\hline
 6 & 8 - 9 & 7 & 20\\\hline
 7 & 9 - 10 & 7 & 20\\\hline
 8 & 10 - 11 & 7 & 20\\\hline
 9 & 11 -12 & 7 & 20\\\hline
 10 & 12 -13 & 7 & 20\\\hline
 11 & 13 - 14 & 7 & 20\\\hline
 12 & 14 - 15 & 9 & 33\\\hline
 Easter & 15 - 17 & 12 & {}-\\\hline
 13 & 17 - 18 & 7 & 35\\\hline
 14 & 18 - 19 (Leading up to event) & 9 & 35\\\hline
 After & 19 - 22 & 21 & 65\\\hline
 Total: & & 91 & 383\\\hline
\end{longtable}


\pagebreak
\subsubsection{Actual Work}
Table~\ref{table:actualWork} shows the actual sprints and work done. The hours
are for each person, during that sprint.
\begin{longtable}{|l|l|l|l|}
\caption{Actual work} \label{table:actualWork} \\
\hline
 Sprint & Week & Days & Hours \\\hline
 1 & 3-4 & 7 & 15\\\hline
 2 & 4-5  & 7  & 15\\\hline
 3 & 5-6 & 7  & 20\\\hline
 4 & 6-7 & 7 & 20\\\hline
 5 & 7-8 & 7 & 27\\\hline
 6 & 8-9  & 7  & 31\\\hline
 7 & 10-11 & 7  & 35\\\hline
 8 & 11-12 & 7 & 30\\\hline
 9 & 12-13 & 7  & 30\\\hline
 10 & 14-15 & 9 & 40\\\hline
 11 & 15-17 \ (starting 16.04, ending 26.04, easter) & 10 & 90\\\hline
 12 & 18-19 & 6 & 35\\\hline
 After & 19-22 & 21 & 80\\\hline
 Total & & 100 & 468\\\hline
\end{longtable}
