\chapter{Introduction}


\section{About the course}

Our group and assignment has been delegated as part of the course
IT2901: {\textquotedblleft}Informatics Project II{\textquotedblright}
at NTNU. The work covers 15 course credits, equivalent to a 50\% work
position for one academic semester. IT2901 is offered only to those
that are enrolled on the NTNU{\textquoteright}s informatics BSc
programme.


The primary purpose of the course is to let students apply their
knowledge from other courses. This is rendered through a project for a
real customer. The students have to communicate independently with
their client, and deliver a software product that answers the
client{\textquoteright}s needs.


Grades are based on the satisfaction of the customers and an evaluation
of the development process. The latter will be reviewed through written
reports and timesheets, as provided in this document. Furthermore, it
is important that students have met the given deadlines and documented
their work in a structured manner.

\section{The Group}

The team consists of six members. All the members of the group are
completing their BSc degree in Computer Science from NTNU in 2014. We
had prior experience working together, and knew each other well. With
many shared courses and similar interests, the team are all at a
somewhat similar level of competence. However, we have different areas
of expertise, and exploiting this has been a key to success on previous
occasions. For a detailed description of each member, see the listing
below.



\subsection*{Anders Sildnes}

Throughout his BSc, Anders has been taking courses related to algorithms
and program security. Apart from his studies, he is developing for
Engineers without Borders NTNU and spending time with open-source
projects and other Linux tools.

\subsection*{Eirik Fosse}

Eirik has a primary interest in artificial intelligence and machine
learning. In the course of his bachelor{\textquoteright}s degree
he{\textquoteright}s focused on programming, mathematics, and
evolutionary simulation.


\subsection*{Filip Fjuk Egge}

While achieving his degree, Filip has taken courses focused on a path
related to system development and security. He has a varied education
and knowledge on different aspects of computer science. 


\subsection*{Haakon Konrad William Aasebø}

Haakon has selected disciplines related to mathematics and algorithms.
Apart from being a student at NTNU he is playing football at NTNUI in
the third division.

\subsection*{Håkon Gimnes Kaurel}

During his time at NTNU, Håkon has been keeping a primary focus on
courses related to programming and the intersection between hardware
and software. He{\textquoteright}s also got experience as an app
developer, and has extensive knowledge of the GNU/Linux operating
system.

\subsection*{Tino Lazreg}

Tino has been taking courses related to different aspects of software
engineering, like programming, system architecture, human-machine
interaction. Besides doing a BSc, Tino also works as a student
assistant in a human-machine interaction course on NTNU. 


\section{About the Customer}

Our customer is IDI Open. They are responsible for the annual
programming contest mentioned in %
%Nevnes ikke i 1.2, eller er nummereringen feil?
%
%feil nummerering.
1.2. Christian Chavez is our main contact for the project, but his two
colleagues, Christian Neverdal Jonassen and Finn Inderhaug Holme, were
also available for questions. They are all students of computer science
at NTNU.


\section{About the Contest}

IDI Open is a programming contest where teams of up to three people meet
and solve programming problems of various difficulty. The contest lasts
five hours, and the objective is to solve as many problems as possible.
The contest is open for all types of programmers, from students of all
grades to professors and other professionals from the IT industry.
Various prizes are given to the teams based on their performance. %
%gjentagelse
The main objective, however, is to solve the highest amount of problems
in the shortest amount of time.


There are usually 8-12 problems in a contest. They are of varying
difficulty, ranging from easy to hard. To make the competition fun for
everyone, there are typically some problems that are easy enough even
for novice programmers to handle.

\section{Stakeholders}

Our stakeholder fall into two different categories: the ones involved in
the competition, and those involved in the course.
\subsection{Course}


\subsubsection{Supervisor}

The supervisor{\textquoteright}s job consists of guiding and helping us
through this project. This aid was primarily focused on the development
process and the writing of this report. The supervisor %
%høres litt feil ut
will want to ensure that the developers communicate properly and have a
structured approach to developing the end product. To verify this, we
have had bi weekly status reports delivered to the supervisor, as well
as regular meetings.

\subsubsection{Examiner}

The examiner(s) is responsible for determining our final grade. Opposed
to most of the other stakeholders, we{\textquoteright}ve got little to
no on-going communication with the examiner. However, the examiner has
got access to all the documents the supervisor has got access to.

\subsection{Product}


\subsubsection{IDI Open}

The project{\textquoteright}s primary stakeholders. They are the host of
the competition in which our product was used. Their inclusion in this
product comprised all aspects of our project.

\subsubsection{Judges}

The judges are hired by %
%IDI Open?
IDI NTNU to supervise the competition, service contestants and create
problem sets. They will rely on our end product achieve the mentioned
tasks. Throughout the process they have %
%hva har de rapportert
reported to our customers, IDI Open. Naturally, the judges are important
to the contest, so it is important that they are satisfied with the
software they have to use.

\subsubsection{Developers}

The developers are responsible for satisfying all other parties. Similar
to the customer, our involvement in this project is total.


\subsubsection{Maintainers}

As IDI Open is an annually recurring event, our end product, if
successful, will be used for many years in the future. At a point, we
assume the code will need to be replaced or modified. Assumably, there
will be another developer team to do this. As such, the quality of of
our product will impact them.

\subsubsection{Sponsors}

Different companies sponsor each contest. In exchange for money and
services, the sponsors get exposure through ads on the website and get
to give a short presentation during the awards ceremony. Naturally, the
sponsors want to associate their name with a successful product.
Therefore, the sponsors rely on that contests are successful - this is
heavily based on our product.

\subsubsection{Contestants}

The actions of contestants are all through our software; our product
will be their medium to take part in IDIOpen. Reliability and usability
is key to keep the contestants happy. For example system flaws
affecting the contest results will not be tolerated. %
%Hva slags feedback, hvordan brukte vi dette
Our customers also received feedback from the contestants. Thus, how
satisfied the contestants are impacts the developer{\textquoteright}s
evaluation.

\subsection{Goals}

The goal of the project is to upgrade and improve the existing system
used in IDI Open. %
%We were given sole responsibility for our project; no other team or organization of developers has had responsibility for our solution.
%
%Liker din setning bedre
The project is not a prototype wanted by a consulting firm, but a system
that will be used by actual end users. This gave us inspiration to do
the best we could, and to give the customer something both we and they
could be proud of for many years to come. And if the product is good
enough it would hopefully also be used in larger programming
competitions, maybe even international ones.

