
\chapter{Installation guide}
This is the complete install guide for GentleIDI. The guide will assume
that the reader is has got some basic linux skills. You should be
capable of installing packages by means of a package-manager like apt,
yum etc. \ 


\bigskip

Though GentleIDI is not tightly linked with any specific linux distro,
this guide assumes that you{\textquoteright}re using Ubuntu Server
14.04. This is the only distro on which the system has been tested
thoroughly at the time of writing. 


\bigskip

GentleIDI is in many ways a straightforward django-based website, and
hence there are a lot of possible setups to choose from. This guide is
inspired by a guide written by Michal Karzy\'nski, and will guide you
through the steps of setting up the system using a combination of
gunicorn and nginx. \ 

\section{Creating your users}
Running a website as a user with root privileges or anything of the sort
is far from recommended. Therefore you are recommended to create a new
user and a new usergroup. The names of both the group and the user can
be chosen as you please, but the rest of the guide will stick to using
a user called gentleidi and a group named webapps.




\$ sudo mkdir -p /webapps/gentleidi

\$ sudo groupadd -{}-system webapps. \newline
\$ sudo useradd -{}-system -{}-gid webapps -{}-home /webapps/gentleidi
gentleidi\newline
\$ sudo chown gentleidi:webapps /webapps/gentleidi/

Now you have a user named gentleidi which is a member of the usergroup
webapps, and whose home directory is /webapps/gentleidi.


\bigskip

In addition to the user we just created, we need another user,
specifically used to run the untrusted software submitted by the
contestants. GentleIDI assumes that this user is named gentlemember.
However, changing this value in the source is no complicated matter. 


\bigskip

\$sudo useradd -{}-system gentlemember


\bigskip

The system needs to be able to execute commands both as gentleidi and
gentlemember. As the web server runs as gentleidi we need to make sure
that gentleidi can execute commands as gentlemember. Add the following
line to your sudoers file.


\bigskip

gentleidi ALL=(gentlemember) NOPASSWD:ALL


\bigskip

If you don{\textquoteright}t know how to edit your sudoers, to open the
sudoers file in a text editor simply type the following command:


\bigskip

\$sudo visudo


\bigskip

Now we{\textquoteright}ve got two users, one capable of executing
commands as the other. What we want to do now is to ensure that
gentlemember is unable to communicate via network. This is done by
applying two rather straightforward iptable rules.


\bigskip

\$ sudo iptables -A OUTPUT -m owner -{}-uid-owner gentlemember -j LOG

\$ sudo iptables -A OUTPUT -m owner -{}-uid-owner gentlemember -j REJECT


\bigskip

Though this will restrict the user{\textquoteright}s network access, be
aware of software installed on your system which is capable of
switching to another user. 


\bigskip

\section{Setting up the environment}

Due to a lot of strict changes made in python versions, a lot of
libraries do not work across different versions of python. This leaves
python in a situation where program A might need python to be version X
and program B might need python to be version Y. So, what do you do?
You setup a virtual environment.


\bigskip

Virtual environments is a way of setting up separate python setups for
different sets of programs. 


\bigskip

What we want to do is to turn the home directory of the gentleidi user
into a virtual environment. 


\bigskip

\$ sudo apt-get install python-virtualenv

\$ sudo su gentleidi

\$ virtualenv \~{}/env


\bigskip

Now that you{\textquoteright}ve got a virtual environment you can start
filling it with something useful, like the content of the
project{\textquoteright}s git repo. 


\bigskip

%
%er det bedre å ha git repoet direkte?
%
%Hva mener du? Nå blir jo repoet direkte? Kopierer jo bare innholdet i IDIOpen.
%
%å ha det i webapps/gentleidi istedet for å kopiere
%
%Jeg kopierer jo alt innholdet i repoet inn der, også .git/. Så jeg lager jo bare et nytt repo i /webapps/gentleidi på sett og vis. Hovedforskjellen er jo bare at virtualenv forblir der også.
%
%På virtualbox la jeg virtual env i /webapps/gentleidi/env/ og clonet inn i IDIOpen mappen
%
%Noe sånt?
\$ cp -r /path/to/repo/IDIOpen/ /webapps/gentleidi/


\bigskip

Please note that you only need the wsgi folder from the repo, however,
updating is a lot easier when all you{\textquoteright}ve got to do is
pull the latest version directly using git. The downside is that you
could possibly end up committing your production system configuration
files etc to the repo. However, we{\textquoteright}re going to assume
that you will not be developing directly in your production system, and
thereby avoid the hazard. 


\bigskip

Before leaving this step, ensure that the files in /webapps/gentleidi
has got the correct file permissions. 


\bigskip

\$ sudo chown -R gentleidi:webapps /webapps/gentleidi


\bigskip

\section{Installing required packages}
Now it{\textquoteright}s time to start making sure that
you{\textquoteright}ve got the packages you need to run GentleIDI. 


\bigskip

\$ sudo apt-get install git nginx libmysqlclient-dev python-dev


\bigskip

You might already have most of these packages, however, better safe than
sorry. 


\bigskip

The next thing you need to do before continuing is to log in as
gentleidi and activate your newly created virtual environment.


\bigskip

\$ sudo su gentleidi

\$ source \~{}/env/bin/activate


\bigskip

Installing the required python packages via pip is easily done. In the
project root directory there{\textquoteright}s a file named
requirements.txt. This file is simply a list of required packages, to
install them simply execute the following:


\bigskip

\$ pip install -r requirements.txt


\bigskip

\section{Database}
GentleIDI needs a database to store its data. This guide will show you
how to setup GentleIDI with a mysql database server, however, if you
feel like using mariaDB, postgresql, or even SQLite, then please do.
Any database server supported by Django is supported by GentleIDI. 


\bigskip

Naturally you don{\textquoteright}t need to install the database server
on the same host as the web server, that{\textquoteright}s what
we{\textquoteright}ll do for now. 


\bigskip

\$ sudo apt-get install mysql-server


\bigskip

Now what we need to do is to create a database and a mysql user that
GentleIDI can use. During the install process you were required to set
a root password for the mysql-server. Login as root and perform the
following commands:


\bigskip

{\textgreater} CREATE USER
{\textquoteleft}gentledb{\textquoteright}@{\textquoteright}localhost{\textquoteright}
IDENTIFIED BY {\textquoteleft}password{\textquoteright};

{\textgreater} GRANT ALL PRIVILEGES ON * . * TO
{\textquotesingle}newuser{\textquotesingle}@{\textquotesingle}localhost{\textquotesingle};

{\textgreater} FLUSH PRIVILEGES;

{\textgreater} CREATE DATABASE gentleidi CHARACTER SET uft8 COLLATE
utf8\_general\_ci;


\bigskip

Remember to replace {\textquoteleft}gentledb{\textquoteright} and
{\textquoteleft}password{\textquoteright} with a suitable username and
password. \newline


Now you need to ensure that GentleIDI uses your newly created database.
Edit the DATABASES entry in IDIOpen/wsgi/openshift/settings.py


\bigskip

if MYSQL:
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

130 \ \ \ \ DATABASES = \{
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

131 \ \ \ \ \ \ \ \ \ {\textquotesingle}default{\textquotesingle}: \{
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

132 \ \ \ \ \ \ \ \ \ \ \ \ \ {\textquotesingle}ENGINE{\textquotesingle}
\ \ : {\textquotesingle}django.db.backends.mysql{\textquotesingle},
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

133 \ \ \ \ \ \ \ \ \ \ \ \ \ {\textquotesingle}NAME{\textquotesingle} :
{\textquotesingle}gentleidi{\textquotesingle},
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

134 \ \ \ \ \ \ \ \ \ \ \ \ \ {\textquotesingle}USER{\textquotesingle} :
{\textquotesingle}gentledb{\textquotesingle}, \ \ \ \ \ \ \ \ \ \ \ \ 

135 \ \ \ \ \ \ \ \ \ {\textquotesingle}PASSWORD{\textquotesingle} :
{\textquotesingle}password{\textquotesingle},
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

136 \ \ \ \ \ \ \ \ \ {\textquotesingle}HOST{\textquotesingle} :
{\textquotesingle}localhost{\textquotesingle},
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

137 \ \ \ \ \ \ \ \ \ {\textquotesingle}PORT{\textquotesingle} :
{\textquotesingle}3306{\textquotesingle},
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

138 \ \ \ \ \ \ \ \ \ \}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

139 \ \ \ \ \}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 


\bigskip

In order to make sure that the database is working properly, login as
gentleidi, activate your environment and synchronize
GentleIDI{\textquoteright}s database.


\bigskip

\$ sudo su gentleidi

\$ source \~{}/env/bin/activate

\$ python \~{}/IDIOpen/wsgi/manage.py syncdb

\$ python \~{}/IDIOpen/wsgi/manage.py migrate


\bigskip

If this command terminates properly, then your database should be good
to go. In fact you should be able to run GentleIDI on a development
server at this point. But first, you need to create a admin account.To
do so, simply execute the following:


\bigskip

\$ python \~{}/IDIOpen/wsgi/openshift/manage.py createsuperuser


\bigskip

To start the development server run:


\bigskip

\$ python \~{}/IDIOpen/wsgi/openshift/manage.py runserver 


\bigskip

You should now have a working website running on port 8000. However, you
have no execution nodes available to evaluate submissions, and
you{\textquoteright}re using Django{\textquoteright}s development
server, which scales horribly. 


\bigskip

\section{Gunicorn}
Now it{\textquoteright}s time to install replace the Django development
server with a proper application server, gunicorn. Member to be logged
in as gentleidi, and to activate your environment before doing this. 


\bigskip

\$ pip install gunicorn


\bigskip

Now we need a script that launches gunicorn and GentleIDI appropriately.


\#!/bin/bash


\bigskip

NAME={\textquotedbl}GentleIDI{\textquotedbl}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \#
Name of the application

DJANGODIR=/webapps/gentleidi/IDIOpen/wsgi/ \ \ \ \ \ \ \ \ \# Django
project directory

SOCKFILE=/webapps/gentleidi/run/gunicorn.sock \ \ \ \# we will
communicate using this unix socket

USER=gentleidi
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \#
the user to run as

GROUP=webapps
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \# the
group to run as

NUM\_WORKERS=3 \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \# how many
worker processes should Gunicorn spawn

DJANGO\_SETTINGS\_MODULE=openshift.settings \ \ \ \ \ \ \ \ \ \ \ \ \#
which settings file should Django use

DJANGO\_WSGI\_MODULE=openshift.wsgi
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \# WSGI module name


\bigskip

echo {\textquotedbl}Starting \$NAME as
{\textasciigrave}whoami{\textasciigrave}{\textquotedbl}


\bigskip

\# Activate the virtual environment

cd \$DJANGODIR

source /webapps/gentleidi/env/bin/activate

export DJANGO\_SETTINGS\_MODULE=\$DJANGO\_SETTINGS\_MODULE

export PYTHONPATH=\$DJANGODIR:\$PYTHONPATH


\bigskip

\# Create the run directory if it doesn{\textquotesingle}t exist

RUNDIR=\$(dirname \$SOCKFILE)

test -d \$RUNDIR {\textbar}{\textbar} mkdir -p \$RUNDIR


\bigskip

\# Start your Django Unicorn

\# Programs meant to be run under supervisor should not daemonize
themselves (do not use -{}-daemon)

exec /webapps/gentleidi/env/bin/gunicorn
\$\{DJANGO\_WSGI\_MODULE\}:application {\textbackslash}

\ {}-{}-name \$NAME {\textbackslash}

\ {}-{}-workers \$NUM\_WORKERS {\textbackslash}

\ {}-{}-user=\$USER -{}-group=\$GROUP {\textbackslash}

\ {}-{}-log{}-level=debug {\textbackslash}

\ {}-{}-bind=unix:\$SOCKFILE


\bigskip


\bigskip


\bigskip


\bigskip


\bigskip

Place the contents of the previous page in the following file:
/webapps/gentleidi/env/bin/gunicorn\_start


\bigskip

Make sure that the script is executable:


\bigskip

\$ sudo chmod u+x /webapps/gentleidi/env/bin/gunicorn\_start


\bigskip

\begin{enumerate}
\item \begin{enumerate}
\item nginx
\end{enumerate}
\end{enumerate}
As mentioned previously this setup relies on a combination of gunicorn
and nginx. At this point gunicorn should be working properly, and
it{\textquoteright}s time to setup nginx. 


\bigskip

If you have not already installed nginx, do so now:

\$ sudo apt-get install nginx


\bigskip

Now you need to create an nginx configuration file for gentelidi. 


\bigskip

Store the content found below in the following file :
/etc/nginx/sites-available/gentleidi

upstream hello\_app\_server \{

\ server unix:/webapps/gentleidi/run/gunicorn.sock fail\_timeout=0;

\}


\bigskip

server \{


\bigskip

\ \ \ listen \ \ 80;

\ \ \ server\_name example.com;


\bigskip

\ \ \ client\_max\_body\_size 4G;


\bigskip

\ \ \ access\_log /webapps/gentleidi/logs/nginx{}-access.log;

\ \ \ error\_log /webapps/gentleidi/logs/nginx{}-error.log;

\ \ \ location /static/ \{

\ \ \ \ \ \ \ alias \ \ /webapps/gentleidi/IDIOpen/wsgi/static/;

\ \ \ \}


\bigskip

\ \ \ location /media/ \{

\ \ \ \ \ \ \ alias \ \ \ /webapps/gentleidi/IDIOpen/wsgi/media/;

\ \ \ \}


\bigskip

\ \ \ location / \{

\ \ \ \ \ \ \ proxy\_set\_header X{}-Forwarded{}-For
\$proxy\_add\_x\_forwarded\_for;

\ \ \ \ \ \ \ proxy\_set\_header Host \$http\_host;

\ \ \ \ \ \ \ proxy\_redirect off;


\bigskip

\ \ \ \ \ \ \ if (!-f \$request\_filename) \{

\ \ \ \ \ \ \ \ \ \ \ proxy\_pass http://hello\_app\_server;

\ \ \ \ \ \ \ \ \ \ \ break;

\ \ \ \ \ \ \ \}

\ \ \ \}


\bigskip

\ \ \ \# Error pages

\ \ \ error\_page 500 502 503 504 /500.html;

\ \ \ location = /500.html \{

\ \ \ \ \ \ \ root \ /webapps/gentleidi/IDIOpen/wsgi/static/;

\ \ \ \}

\}


\bigskip

\#EOF


\bigskip

In this configuration nginx is set to log all accesses and errors. These
files need to be created by the following commands:

\$ sudo su gentleidi

\$ mkdir \~{}/logs

\$ touch \~{}/logs/nginx-access.log

\$ touch logs/nginx-error

\$ exit


\bigskip

All you need to do at this point is to enable the nginx site. This is
done simply by creating a symbolic link from the configuration file in
sites-available to sites-enabled.


\bigskip

\$ sudo ln -s /etc/nginx/sites-available/gentleidi
/etc/nginx/sites-enabled/

\$ sudo rm /etc/nginx/sites-enabled/default

\$ sudo service nginx restart


\bigskip

You should now have a working website. All that{\textquoteright}s left
is making management a little easier and adding some execution nodes.


\bigskip

\section{Supervisor}
Supervisor is a utility for defining and managing jobs. In this case
we{\textquoteright}re going to define two jobs, one for managing the
website, and another for managing an execution node. 


\bigskip

You need to create two files to make this happen:

/etc/supervisor/conf.d/gentleidi.conf

[program:gentleidi]

command = /webapps/gentleidi/env/bin/gunicorn\_start

user = gentleidi

stdout\_logfile = /webapps/gentleidi/logs/gunicorn\_supervisor.log

redirect\_stderr = true

\#EOF


\bigskip

/etc/supervisor/conf.d/celery.conf

[program:celery]
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

command=/webapps/gentleidi/env/bin/celery worker -A openshift -l info
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

directory=/webapps/gentleidi/IDIOpen/wsgi
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

environment=PATH={\textquotedbl}/webapps/gentleidi/env/bin:\%(ENV\_PATH)s{\textquotedbl}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

user=gentleidi
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

autostart=true
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

autorestart=true
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 

redirect\_stderr=True \ \ 

\#EOF


\bigskip

Create the log files that you{\textquoteright}ve referenced. 

\$ mkdir /webapps/gentleidi/logs/

\$ touch /webapps/gentleidi/logs/gunicorn\_supervisor.log


\bigskip

Read the newly created configuration files.

\$ sudo supervisorctl reread 

\$ sudo supervisorctl update

\$ sudo supervisorctl restart all


\bigskip


\bigskip

\section{Multiple execution nodes}
The easiest way of setting up multiple execution nodes is to clone the
setup on your web server to other machines and then making minor
changes.


\bigskip

When setting up multiple execution nodes there are two changes that need
to be made. The directory
/webapps/gentleidi/IDIOpen/wsgi/private/submissions needs to be shared
between all the execution nodes. How you decide to make this happen is
up to you. However, sshfs is possibly the easiest solution. Whatever
way you decide to mount the directory on your execution nodes, make
sure that multiple users are allowed to access it, e.g. the
allow\_other option for sshfs.


\bigskip

You also need to make sure that all your execution nodes have access to
the same database. Make sure that the settings.py is not set to
localhost, but rather points to whatever host you decide to use as a
database server. Some configuration of your database server might be
needed in order for it to accept remote connections. mysql servers need
to change the bind-address property in the /etc/mysql/my.cnf to their
actual IP, not localhost(127.0.0.1).


\bigskip

You also need to change the grants for the mysql user in such a way that
it is allowed to connect remotely to the database. 

